### 2020-09-28 results.tex

\section{Main ?? results}

TODO: make a list of images of vectors, e.g.
$\Theta_*(\ddd{\vphi_{\vec{e}}}) = -\ddd{\theta_e}$.

Our general strategry for obtaining a (non-singular) circle pattern
is to start with an assignment $\underline{\theta} = (\theta_\bullet)$
for which we know an extended circle pattern exists
(say from results of \cite{BandS}),
and a path $\gamma$ in $\TTT$ starting at $\underline{\theta}$.
We then attempt to lift $\gamma$ to a path $\tilde{\gamma}$ in $\CCC$,
so that $\tilde{\gamma}$ remains (face) non-singular
(vertex non-singularity is already determined by $\gamma$).

Note that since $\sum_{e\in E} \theta_e
= 2\pi |E| - \sum_{\vec{e} \in \vec{E}} \vphi_{\vec{e}}
= 2\pi |E| - \sum_{f\in F} \Phi_f$,
we see that maintaining face non-singularity of $\tilde{\gamma}$
forces $\sum \theta_e$ to be constant.
We will show that this is the only obstruction on $\gamma$
to the lifting to such $\tilde{\gamma}$.

To that end, let $L$ be the $(|E|-1)$-plane distribution on $\TTT$
tangent to the level sets of $\sum_{e\in E} \theta_e$.
The following proposition proves it (TODO avoid using `it'?) up to first order at a point:

\begin{proposition}
\label{p:point_lift}
Let $c \in \CCC$ be a
non-singular extended circle pattern on a graph
with no bigons or self-loops such that
there is no wide cycle of thin faces.
%\begin{itemize}
	%\item $\Phi_f = 2\pi$ for all $f\in F$;
	%\item there is no wide cycle of thin faces.
	%\item all faces are eithe convex or thin,
		%with at least one convex face;
	%\item every pair of faces is connected by a wide path.
%\end{itemize}
Let $K_c = \ker (\Phi_*|_c : T_c \CCC \to T_{\Phi(c)} \PPP)$.
Then $\Theta_*(K_c) = L_{\Theta(c)}$.

In other words, for any vector $a \in T_{\Theta(c)} \TTT$
with sum of components 0, one can vary $c$ so that its
first order change in $\theta_\bullet$ is $a$,
and also remains face non-singular up to first order.
\end{proposition}

Before we prove this, it is convenient to first prove the following,
which will establish some useful notation:

\begin{lemma}
\label{l:Phi_full}
Let $c \in \CCC$ be as in \prpref{p:point_lift}.
Then $\Phi$ is a submersion in a neighbourhood of $c$.
\end{lemma}

\begin{proof}
We construct vectors $\beta_f \in T_c \CCC$ so that
$\Phi_*(\beta_f) = \ddd{\Phi_f}$.
The vector
\begin{equation}
\label{e:alpha_f}
\alpha_f := \ddd{r_f} - \sum_{\vec{e} \in \del f} 
	\frac{\tan \vphi_{\vec{e}}}{r_f} \ddd{\vphi_{\vec{e}}}
	\in T(\RRR \times \QQQ)
\end{equation}
doesn't change $l_{\vec{e}}$,
so is in $T_c \CCC$.
Intuitively, $\alpha_f$ is like
pulling the center of $C_f$ up off the plane, increasing $r_f$ and decreasing all
$\vphi$'s.
Its pushforward under $\Phi$ is simply
$\Phi_*(\alpha_f) =
	(-\frac{2}{r_f} \sum_{\vec{e}\in \del f} \tan \vphi_{\vec{e}})
	\ddd{\Phi_f}$.

If $f$ is a convex face, then the $\tan \vphi_{\vec{e}}$ are all non-negative,
being 0 if and only if $e$ is short,
and because $\Phi_f = 2\pi$, at least three edges are long.
So $\Phi_*(\alpha_f)$ is a negative multiple of $\ddd{\Phi_f}$;
we choose $\beta$ so that
\begin{equation}
\label{e:beta_f_convex}
\beta_f := \beta \cdot \alpha_f; \;\;\;
\Phi_*(\beta_f) = \ddd{\Phi_f}
\end{equation}

If $f$ is thick but not convex,
let $\vec{e}$ be an edge with largest $\vphi_{\vec{e}}$,
so that $\vphi_{\vec{e}} \geq \pi/2$.
If $\vphi_{\vec{e}} > \pi/2$,
then since $\tan$ is convex, and $f$ has at least two other
long edges,
we see that $\Phi_*(\alpha_f) = -\sum \frac{\tan \vphi_{\vec{e'}}}{r_f} > 0$,
so we can define $\beta_f$ as in \eqnref{e:beta_f_convex}.
If $\vphi_{\vec{e}} = \pi/2$,
we can take
\[
\beta_f = \frac{1}{2} \ddd{\vphi_{\vec{e}}}
\]

If $f$ is thin, we actually have that $\alpha_f \in K_c$,
since the components sum to 0, so we need a different approach.
First suppose $f$ shares an edge $e$ with a thick face $f'$,
with $\vec{e} \in \del f$.
If $\vphi_{\vec{e}} = \pi/2$, we can take
$\beta_f = \frac{1}{2} \ddd{\vphi_{\vec{e}}}$ as before.
Otherwise, we can increase $l_{\vec{e}}$
by varying $\vphi_{\vec{e}}$
while holding $r_f$ constant,
and increase $l_{\cev{e}}$
by increasing $r_{f'}$.
This will affect both $\Phi_f, \Phi_{f'}$,
so we use $\beta_{f'}$ to make $\Phi_{f'}$ constant.
%More explicitly, consider
%\begin{equation}
%\label{e:alpha_e}
%\alpha_e = \frac{1}{2 r_f \cos \vphi_{\vec{e}}} \ddd{\vphi_{\vec{e}}}
%+ \frac{1}{2 r_{f'}\cos \vphi_{\cev{e}}} \ddd{\vphi_{\cev{e}}}
%\end{equation}
%which increases $l_{\vec{e}}, l_{\cev{e}}$ at half unit speed,
%$dl_{\vec{e}}(\alpha_e) = dl_{\cev{e}}(\alpha_e) = 1/2$.
%Then we can take
%\begin{equation}
%\label{e:beta_f_thin}
%\beta_f = r_f \cos \vphi_{\vec{e}}
	%(\alpha_e - \frac{1}{r_{f'} \cos \vphi_{\cev{e}}} \beta_{f'})
%\end{equation}
We can repeat this procedure with $f'$ set to this thin face,
and $f$ set to another thin face adjacent to it, etc.
\end{proof}

\begin{proof}
NEW PROOF NEW PROOF NEW PROOF NEW PROOF NEW PROOF NEW PROOF 
We construct a vector $u_e \in K_c \subseteq T_c \CCC$ for each edge $e$
and show that $\{\Theta_*(u_e)\}_{e\in E}$ spans an
$|E|-1$ dimensional space,
thus must be equal $L_{\Theta_c}$.
$u_e$ will have the following property:
if $\Theta_*(u_e) = \sum_{e' \in E} a^{e'} \ddd{\theta_{e'}}$, then
\begin{itemize}
	\item $a^e = 1$;
	\item $a^{e'} \leq 0$ for all $e' \neq e$;
	%\item $a^{e'} = 0$ for short edges $e'$ (hmm...)
	\item $\sum_{e' \in E} a^{e'} = 0$ (follows directly from $u_e \in K_c$)
\end{itemize}

To illustrate the main idea of the construction,
first suppose that $e$ is an edge between two convex faces $f, f'$
(so $\vec{e} \in \del f, \cev{e} \in \del f'$).
Observe that %$\alpha_e$ from \eqnref{e:alpha_e}
\begin{equation}
\label{e:alpha_e}
\alpha_e = \frac{1}{2 r_f \cos \vphi_{\vec{e}}} \ddd{\vphi_{\vec{e}}}
+ \frac{1}{2 r_{f'}\cos \vphi_{\cev{e}}} \ddd{\vphi_{\cev{e}}}
\end{equation}
increases both $l_{\vec{e}}$ and $l_{\cev{e}}$ at the same rate,
and has the side effect of
increasing $\Phi_f$ and $\Phi_{f'}$.
We can compensate using $\beta_f$ from \eqnref{e:beta_f_convex},
so we consider
\[
w_e = \alpha_e - \frac{1}{\cos \vphi_{\vec{e}}} \beta_f
	- \frac{1}{\cos \vphi_{\cev{e}}} \beta_{f'} \in K_c
\]
%Since $\theta_e = \pi - \vphi_{\vec{e}} - \vphi_{\cev{e}}$,
%we need to show that the coefficients of $\ddd{\vphi_{\vec{e}}},
%\ddd{\vphi_{\cev{e}}}$ sum to 0.
For $\vec{e'} \in \del f \cup \del f' \backslash \{\vec{e},\cev{e}\}$,
the $\ddd{\vphi_{\vec{e'}}}$ component only appears in $\beta_f$
or $\beta_{f'}$, which is positive by construciton
(because $f,f'$ are convex faces);
thus, $d\theta_{e'}(w_e) \geq 0$ for such $e'$, and
is equal 0 if and only if $e'$ is short.
Finally, since $w_e \in K_c$, it leaves the sum $\sum_{e\in E} \theta_e$
constant, so $d\theta_e(w_e) < 0$, so we can take
\begin{equation}
\label{e:u_e}
u_e := (d\theta_e(w_e))^\inv \cdot w_e
\end{equation}

Next we consider the presence of thick faces that are not convex,
but still assume there are no thin faces.
Then a thick face as at most one edge with $\vphi_{\vec{e}} \geq \pi/2$.
Suppose $\vec{e} \in \del f$, and $f$ is not convex.
We want to increase $l_{\vec{e}}$ as before.
If $f$ has no edges with $\vphi_\bullet = \pi/2$,
then we consider
\begin{equation}
\gamma_f = \ddd{\vphi_{\vec{e}}} - \ddd{\vphi_{\vec{e'}}}
\end{equation}
where $e'$ is some other edge of $f$, such that
$\max \{\vphi_{\vec{e}}, \vphi_{\vec{e'}}\} > \pi/2$.
Because $\cos \vphi_{\vec{e}} \neq 0 \cos \vphi_{\vec{e'}}$,
$\gamma_f$ affects both $l_{\vec{e}}$ and $l_{\vec{e'}}$
(increasing or decreasing both simultaneously)
while keeping $\Phi_f$ constant.
This allows us to ``transfer'' the effect of increasing $l_{\vec{e}}$
to $l_{\vec{e'}}$, in the following sense:
if $e'$ is adjacent to another face $f''$ that is convex,
then

END NEW PROOF END NEW PROOF END NEW PROOF END NEW PROOF END NEW PROOF

Let us first construct several other auxiliary vectors
that are not necessarily in $T_c \CCC$.
Generally they keep $\Phi_f$'s constant,
but have an effect on the lengths of oriented edges $l_{\vec{e}}$.
The $u_e$'s will be built from a combination of these.

Let $f$ be a convex face, and $\vec{e} \in \del f$. Define
\begin{equation}
\label{e:eta_e}
\eta_{\vec{e}} =
\frac{1}{2 r_f \cos \vphi_{\vec{e}}} \ddd{\vphi_{\vec{e}}}
- \frac{1}{r_f \cos \vphi_{\vec{e}}} \beta_f
\end{equation}
$\eta_{\vec{e}}$ keeps $\Phi_f$ constant, but increases $l_{\vec{e}}$
at unit speed, i.e. $dl_{\vec{e}}(\eta_{\vec{e}}) = 1$.

Let $f$ be a thick but not convex face.
There is exactly one edge $\vec{e} \in \del f$
with $\vphi_{\vec{e}} \geq \pi/2$.
If $\vphi_{\vec{e}} > \pi/2$, we will define
$\eta_{\vec{e}}$ as in \eqnref{e:eta_e}.
If $\vphi_{\vec{e}} = \pi/2$,
we will define
\begin{equation}
\label{e:eta_e_2}
\eta_{\vec{e}} = \frac{1}{2}\ddd{r_f}
+ (\sum_{\vec{e'} \neq \vec{e}}
	\frac{\sin \vphi_{\vec{e'}}}{2 r_f \cos \vphi_{\vec{e'}}})
	\ddd{\vphi_{\vec{e'}}}
- \sum_{\vec{e'} \neq \vec{e}}
	\frac{\sin \vphi_{\vec{e'}}}{2 r_f \cos \vphi_{\vec{e'}}}
	\ddd{\vphi_{\vec{e'}}}
\end{equation}
Then it is easy to see that
$dl_{\vec{e}}(\eta_{\vec{e}}) = 1$
while
$dl_{\vec{e'}}(\eta_{\vec{e}}) = 0$
for all other $\vec{e'} \neq \vec{e}$,
and $\eta_{\vec{e}}$ keeps $\Phi_f$ constant.

Now let $\vec{e'} \in \del f$ be another edge.
We define
\begin{equation}
\label{e:gamma_e}
\gamma_{\vec{e'}} = \frac{1}{2 r_f \cos{\vphi_{\vec{e'}}}}
	(\ddd{\vphi_{\vec{e'}}} - \ddd{\vphi_{\vec{e}}})
\end{equation}
Then $dl_{\vec{e'}} = 1$,
$dl_{\vec{e}} = -\frac{\cos \vphi_{\vec{e}}}{\cos \vphi_{\vec{e'}}} \geq 0$,
and keeps $\Phi_f$ constant.

Finally, let $f$ be a thin face.
If $\vec{e} \in \del f$ is a short edge,
i.e. $\vphi_{\vec{e}} = 0$,
then we define $\eta_{\vec{e}}$ as in \eqnref{e:eta_e}.
If $\vec{e}$ is long, and $\vphi_{\vec{e}} \neq \pi/2$,
we define $\gamma_{\vec{e}}$ as in \eqnref{e:gamma_e};
if $\vphi_{\vec{e}} = \pi/2$, we define
\begin{equation}
\label{e:xi_e}
\xi_{\vec{e}} = \frac{1}{2}\ddd{r_f}
\end{equation}
Clearly $dl_{\vec{e}}(\xi_{\vec{e}}) = dl_{\vec{e'}}(\xi_{\vec{e}}) = 1$
for the other edge $e'$ with $\vphi_{\vec{e'}} = \pi/2$.
%where $\vec{e'} \in \del f$ is the unique other long edge.

In summary,
$\eta_{\vec{e}}, \gamma_{\vec{e}}, \xi_{\vec{e}}$
always increases $l_{\vec{e}}$ at unit speed,
and keeps $\Phi_{f_{\vec{e}}}$ constant.
$\eta_{\vec{e}}$ does not affect the lenghts of other edges
in the face $f_{\vec{e}}$,
but the other two might affect another edge
- this other edge $\vec{e'}$ is unique,
and $l_{\vec{e}}, l_{\vec{e'}}$ either both increase
or decrease.
Furthermore, $l_{\vec{e'}} \geq l_{\vec{e'}}$.

For any oriented edge $\vec{e}$,
there is exactly one such vector defined for it;
call that vector $\zeta_{\vec{e}}$.

Now we can put these vectors together to construct the desired $u_e$'s.
To illustrate the main idea of the construction,
first suppose that $e$ is an edge between two convex faces $f,f'$.
Then we can consider
\[
w_e = \eta_{\vec{e}} + \eta_{\cev{e}}
\]
Since $dl_{\vec{e}}(w_e) = dl_{\cev{e}}(w_e) = 1$,
and $dl_{\vec{e'}}(w_e) = 0$ for all other edges,
$w_e \in T_c \CCC$;
furthermore, $\eta_\bullet$ does not affect $\Phi_\bullet$,
so $w_e \in K_c$.
For $\vec{e'} \in \del f \cup \del f' \backslash \{\vec{e},\cev{e}\}$,
the $\ddd{\vphi_{\vec{e'}}}$ component only appears in $\beta_f$
or $\beta_{f'}$, which is positive by construciton
(because $f,f'$ are convex faces);
thus, $d\theta_{e'}(w_e) \geq 0$ for such $e'$, and
is equal 0 if and only if $e'$ is short.
Finally, since $w_e \in K_c$, it leaves the sum $\sum_{e\in E} \theta_e$
constant, so $d\theta_e(w_e) < 0$, so we can take
\begin{equation}
\label{e:u_e}
u_e := (d\theta_e(w_e))^\inv \cdot w_e
\end{equation}

Now consider an arbitrary edge $e$ between faces $f,f'$.
We build $u_e$ inductively, beginning with
$w_e = \zeta_{\vec{e}} + \zeta_{\cev{e}}$,
$F' = \{f, f'\}$, and $E' = \{e\}$.
If $w_e$ also affects $l_{\vec{e'}}$ for some edge
$\vec{e'} \in \del F' \backslash\{\vec{e''},\cev{e''} | e''\in E'\}$,
we add an appriopriate positive multiple
of $\zeta_{\cev{e'}}$ to $w_e$,
add $f_{\cev{e'}}$ to $F'$,
and add $e'$ to $E'$.
Recall that when $\zeta_{\vec{e'}}$ affects
the length of another edge $e''$,
we always have $l_{\vec{e''}} \geq l_{\vec{e'}}$.
Thus by the condition of non-existence of wide cycles of thin faces,
this process must terminate.

Now we need to check that

\end{proof}

\begin{proof}[Proof of \prpref{p:point_lift}]
We construct a vector $u_e \in K_c \subseteq T_c \CCC$ for each edge $e$
and show that $\{\Theta_*(u_e)\}_{e\in E}$ spans an
$|E|-1$ dimensional space,
thus must be equal $L_{\Theta_c}$.

For long edges $e$, $u_e$ will have the following property:
if $\Theta_*(u_e) = \sum_{e' \in E} a^{e'} \ddd{\theta_{e'}}$, then
\begin{itemize}
	\item $a^e = 1$;
	\item $a^{e'} \leq 0$ for all $e' \neq e$;
	\item $a^{e'} = 0$ for short edges $e'$.
	\item $\sum_{e' \in E} a^{e'} = 0$ (follows directly from $u_e \in K_c$)
\end{itemize}
Furthermore, these $u_e$'s collectively satisfy the following
connectivity property:
consider the graph $G_\llong$ whose vertex set is the set of long edges,
and we connect two long edges $e,e'$ by an edge if there is some $e"$ such that
$a^e, a^{e'}$ are both nonzero in $\Theta_*(u_{e"})$;
then $G_\llong$ is connected.

In addition, for short edges $e$,
$u_e$ have the following property: again writing
$\Theta_*(u_e) = \sum_{e' \in E} a^{e'} \ddd{\theta_{e'}}$,
one has
\begin{itemize}
	\item $a^e = 1$;
	\item $a^{e'} = 0$ for other short edges $e'$.
	\item $\sum_{e' \in E} a^{e'} = 0$ (follows directly from $u_e \in K_c$)
\end{itemize}

Let us first suppose we have constructed such $u_e$,
and show that these properties ensure that
$\{\Theta_*(u_e)\}_{e\in E}$ spans an $|E|-1$ dimensional space.
This is a simple exercise in linear algebra,
but we show it for completeness.

Put the $\Theta_*(u_e)$'s into a $E \times E$ matrix,
denoted $M$,
so that the $e$-th row corresponds to $u_e$.
By virtue of $u_e \in K_c$, we have that $(1\; 1\; \cdots \; 1)^T$
is in the null space of $M$;
our goal is to show that it spans the null space.

Suppose $b = (b_e)^T$ is in the null space of $M$.
Reorder $E$, if necessary, so that long edges come before
short edges.
Then by the last property of $u_e$'s for long edges,
the top right block of long-by-short entries are 0.
Thus the vector $b_\llong := (b_e)_{e \in \llong}^T$ is in the null space
of the submatrix $M_\llong$ of long-by-long entries.

Let $|b_e|$ be the largest among components of $b_\llong$;
rescale $b$ so that $b_e = 1$.
The $e$-th component of $M_\llong b_\llong$
is $1 - \sum_{e' \neq e} a^{e'} b_{e'}$,
where $a^{e'}$ are the components of $\Theta_*(u_e)$.
Since $\sum_{e' \neq e} |a^{e'}| = 1$, this can be 0 if and only if
for all $e'$ with $a^{e'} < 0$, we have exactly $b_{e'} = 1$.
By connectedness of $G_\llong$, this implies
$b_\llong = (1 \; \cdots \; 1)^T$.

Returning to the full $M$, it is now easy to see that
for a short edge $e$,
by looking at the $e$-th row of $M$,
we have $b_e = \sum a^{e'} = 1$,
so we are done.

%\footnote{
%This is a well-known argument/result about graph Laplacians:
%the first two properties of $u_e$'s ensure that
%$-M_\llong^T$ is essentially a graph Laplacian for $G_\llong$,
%so has rank = (\#vertices) - (\#components).}

Now we construct the $u_e$'s.
First suppose that $e$ is a long edge between two convex faces $f, f'$
(so $\vec{e} \in \del f, \cev{e} \in \del f'$).
Observe that $\alpha_e$ from \eqnref{e:alpha_e}
increases $\Phi_f$ and $\Phi_{f'}$,
and we can compensate using $\beta_f$ from \eqnref{e:beta_f_convex},
so we consider
\[
w_e = \alpha_e - \frac{1}{\cos \vphi_{\vec{e}}} \beta_f
	- \frac{1}{\cos \vphi_{\cev{e}}} \beta_{f'} \in K_c
\]
%Since $\theta_e = \pi - \vphi_{\vec{e}} - \vphi_{\cev{e}}$,
%we need to show that the coefficients of $\ddd{\vphi_{\vec{e}}},
%\ddd{\vphi_{\cev{e}}}$ sum to 0.
For $\vec{e'} \in \del f \cup \del f' \backslash \{\vec{e},\cev{e}\}$,
the $\ddd{\vphi_{\vec{e'}}}$ component only appears in $\beta_f$
or $\beta_{f'}$, which is positive by construciton
(this requires $f,f'$ to be convex faces);
thus, $d\theta_{e'}(w_e) \geq 0$ for such $e'$, and
is equal 0 if and only if $e'$ is short.
Finally, since $w_e \in K_c$, it leaves the sum $\sum_{e\in E} \theta_e$
constant, so $d\theta_e(w_e) < 0$, so we can take
\begin{equation}
\label{e:u_e}
u_e := (d\theta_e(w_e))^\inv \cdot w_e
\end{equation}

In general, we consider a wide path of faces $f_0, \ldots, f_n$,
such that all faces except $f_0$ and $f_n$ are thin,
and $f_i$ are distinct except possibly $f_0 = f_n$.
\footnote{Picture a closed accordian.}
Let $e_i$ be the long edge between $f_i$ and $f_{i+1}$,
and orient so that $\vec{e_i} \in \del f_i$, $\cev{e_i} \in \del f_{i+1}$.

We will construct $u_{e_i}$ for $i = 1,\ldots,n$.

We want to increase the lengths of $e_i$'s all together.
Let $E' = \{\vec{e_i}, \cev{e_i} | i=0,1,\ldots,n-1\}$.
Consider
\[
x :=
\sum_{i=0,\ldots,n-1} \alpha_{e_i}
=
\sum_{\vec{e} \in E'}
\frac{1}{2 r_{f_{\vec{e}}} \cos \vphi_{\vec{e}}} \ddd{\vphi_{\vec{e}}}
%\sum_{\substack{i = 0,\ldots, n-1 \\ \vec{e} = \vec{e_i}, \cev{e_i} }}
%\frac{1}{2 r_{f_0} \cos \vphi_{\vec{e_0}}} \ddd{\vphi_{\vec{e_0}}}
%+
%\frac{1}{2 r_{f_n} \cos \vphi_{\cev{e_{n-1}}}} \ddd{\vphi_{\cev{e_{n-1}}}}
%+
\]
with $\alpha_e$ from \eqnref{e:alpha_e}.
It is clear that $x$ increases all oriented $e_i$ lengths
$l_{\vec{e}}$ equally fast,
i.e. $dl_{\vec{e_i}}(x) = dl_{\cev{e_i}}(x) = 1/2$ for all $i$,
so $x\in T_c \CCC$.
It is also clear that for the thin faces $f_i$,
$x$ makes one of the angles $\vphi_{\cev{e_{i-1}}}, \vphi_{\vec{e_i}}$
increase and the other decrease, both at the same rate,
thus $d\Phi_{f_i}(x) = 0$.
However, the vector $x$ increase $\Phi_{f_0}$ and $\Phi_{f_n}$,
so it is compensated using $\beta_f$'s, so that
\[
y := x - \frac{1}{r_{f_0} \cos \vphi_{\vec{e_0}}} \beta_{f_0}
- \frac{1}{r_{f_n} \cos \vphi_{\cev{e_{n-1}}}} \beta_{f_n}
\in K_c
\]
Before we modify $y$ to give the desired $u_{e_i}$,
we note that for $\vec{e} \in (\cup \del f_i) \backslash E'$,
the $\vphi_{\vec{e}}$-component in $y$ is non-positive,
and is 0 if and only if $e$ is short;
so $\Theta_*(-y)$ already satisfies three of the four properties
desired in $\Theta_*(u_e)$.

Recall that for thin face $f$,
$\alpha_f$ from \eqnref{e:alpha_f} is in $K_c$,
and we have
\[
\Theta_*(\frac{r_{f_i}}{\tan \vphi_{\vec{e_i}}} \alpha_{f_i}) = 
	\ddd{\theta_{e_i}} - \ddd{\theta_{e_{i-1}}}
\]
Write $\bar{\alpha}_{f_i} = \frac{r_{f_i}}{\tan \vphi_{\vec{e_i}}} \alpha_{f_i}$.
Then it is easy to see that using $\bar{\alpha}_{f_i}$'s, we can
``push around" the $\theta_{e_i}$-components in $\Theta_*(y)$
so only one is nonzero.
More precisely, if
$\Theta_*(y) = \sum_e y^e \ddd{\theta_e}$, the image under $\Theta_*$
of the vector
\[
\bar{u}_{e_0} := y - y^{e_{n-1}} \bar{\alpha}_{f_{n-1}}
- (y^{e_{n-2}} + y^{e_{n-1}}) \bar{\alpha}_{f_{n-2}}
- \ldots
- (y^{e_1} + \ldots + y^{e_{n-1}}) \bar{\alpha}_{f_1}
\]
would have $\theta_{e_i}$-components 0 for all $i$ except 0;
for $i=0$, the $\theta_{e_0}$-component is now $y^{e_0} + \ldots + y^{e_{n-1}}$,
which is nonzero because $\sum_{e'\in E} y^{e'} = 0$
and $\sum_{e' \in E \backslash E'} y^{e'} > 0$.
Therefore, we can take
\[
u_{e_0} := (d\theta_{e_0}(\bar{u}_{e_0}))^\inv \cdot \bar{u}_{e_0};
\;\;\;
u_{e_i} = u_{e_{i-1}} + \bar{\alpha}_{f_i}
\]

We still need to show the collective connectivity property of the $u_e$'s.
It is easy to see that the long edges of a convex face
are connected in $G_\llong$
(for example, $u_{e_0}$ defined above clearly does it for
$f_0$).
Furthermore, all the $\Theta_*(u_{e_i})$ have most components equal,
so $e_i$ are in the same connected component of $G_\llong$.
Thus connectivity of $G_\llong$ follows from the
wide-path-connectivity of faces.

Finally, for short edges $e$, we simply define $u_e$
as in \eqnref{e:u_e}
(here we do not worry about whether faces are thin or convex,
as the non-positivity condition of $a^{e'}$ is absent).
\end{proof}

\begin{proposition}
\label{p:nghd_lift}
Let $c\in \CCC$ be as in \prpref{p:point_lift}.
Given a smooth path $\gamma: [0,\infty) \to \TTT$
starting at $\gamma(0) = \Theta(c)$
and has constant value $(\sum \theta_e) \circ \gamma$,
there exists $\veps > 0$ and a lift $\tilde{\gamma} :[0, \veps] \to \CCC$
of $\gamma|_{[0,\veps]}$ along $\Theta$
that starts at $\tilde{\gamma}(0) = c$
and has constant value $\Phi \circ \tilde{\gamma}$.
\end{proposition}

\begin{proof}
By \lemref{l:Phi_full}, $\Phi$ is a submersion at $c$,
hence is a submersion in a neighborhood $U \in \CCC$ of $c$.
Thus, the kernels
$K_{c'} = \ker(\Phi_*|{c'} : T_{c'} \CCC \to T_{\Phi(c')} \PPP)$
form an $|E|$-plane distribution $K \subset T\CCC|_U$ over $U$.
By \prpref{p:point_lift}, the bundle map
$\Theta_*|_K : K \to L$ is full rank at $c$,
hence it is full in a possibly smaller neighborhood,
which we redefine $U$ to be.

Since $\sum_{e \in E} \theta_e = 2\pi |E| - \sum_{f\in F} \Phi_f$,
the fullness of $\Phi_*$ also shows that
$\sum \theta$ has no critical points in $U$;
for $s$ in some small interval around $(\sum \theta)(c)$,
denote, by overloading notation,
$L_s = (\sum \theta)^\inv (s)$,
$U_s = U \cap \Theta^\inv(L_s)$.
We may consider $K$ as a family of $|E|$-plane distributions
$K_s$ over $U_s$.

The vector $\rho := \sum_{f\in F} r_f \ddd{r_f}$ corresponds to
scaling all radii by the same factor, and is clearly in $T\CCC$.
It is obviously mapped to zero under $\Phi_*$ and $\Theta_*$,
so $\Span\{\rho\} = \ker(\Theta_*|_K) \subset K$.
Thus, over $U$, we may split
$K = \Span\{\rho\} \oplus K'$
(say as orthogonal decomposition w.r.t. some smooth metric on $\CCC$),
so that $\Theta_*|{c'} : K'|_{c'} \simeq L_{\Theta(c')}$.

In other words, $K'|_{U_s}$ is a horizontal $(|E|-1)$-plane distribution
that determines an Ehresmann connection of the
fibre bundle $U_s \to L_s$ (after being appropriately restricted).
Therefore, a short path $\gamma$ starting at $\Theta(c)$
with constant $\sum \theta$, i.e. a path in $L_s$, $s ={(\sum \theta)(c)}$,
can be lifted to a path $\tilde{\gamma}$ in $U_s$,
so that $\tilde{\gamma}$ is always tangent to $K'$,
hence $\Phi(\tilde{\gamma})$ is constant.
\end{proof}

%===============================================================================

TODO
Note: no face can have its edges glued together; prevented by fact
of alternating links

TODO just use Feasible Flow Theorem (see BS)